---
title: "Assignment 2 - Language Development in ASD - Part 1 - Explaining development, Study Group 7"
author: "Kristian Severin, Lasse Hansen, Nikolaj Munch & Sarah Nielsen"
date: "16/09/2020"
output: html_document
---

# Assignment 2 - Part 1

### Loading the relevant libraries

```{r Load Libraries, include = FALSE}
pacman::p_load(tidyverse, lme4, MuMIn, ggpubr, nlme, lmerTest, sjstats, ggplot2, gganimate, gifski)
```

### Loading data

```{r Load Data, include = FALSE}
df <- read_csv("data.csv")
df$X1 <- NULL
```

## Describe your sample (n, age, gender, clinical and cognitive features of the two groups) and critically assess whether the groups (ASD and TD) are balanced "CODE PART"

```{r descriptive stats, include = FALSE}
data %>% count(SUBJ)
# 66 participants

data %>% count(SUBJ) %>% 
  filter(n == 6)
# 58 participants completed all 6 visits

# Creating dataframe with those subjects who completed all visits
alltrials <- data %>% 
  count(SUBJ) %>% 
  filter(n == 6) %>% 
  left_join(data, "SUBJ")

alltrials %>% filter(VISIT == 6) %>% 
  count(Diagnosis)
# 27 ASD children and 31 TD children

alltrials %>% filter(VISIT == 6) %>% 
  count(Gender)
# 10 females and 48 males

alltrials %>% filter(VISIT == 6) %>% 
  filter(Diagnosis == "ASD") %>% 
  count(Gender)
# Of ASD children 4 were F and 23 were M

alltrials %>% filter(VISIT == 6) %>% 
  filter(Diagnosis == "TD") %>% 
  count(Gender)
# Of TD children 6 were F and 25 were M

alltrials %>% 
  group_by(Diagnosis) %>% 
  summarise(mean = mean(Age, na.rm = T))
# Mean age of TD children 30.77 months and 43.45 of ASD children

alltrials %>% 
  group_by(Diagnosis) %>% 
  summarise(mean = mean(MullenRaw, na.rm = T))
# Mean MSEL of TD children was 36.34 and 33.45 in ASD children

alltrials %>% 
  group_by(Diagnosis) %>% 
  summarise(mean = mean(ExpressiveLangRaw, na.rm = T))
# Mean MSEL of TD children was 29.87 and mean MSEL of ASD children was 23.13

alltrials %>% 
  group_by(Diagnosis) %>% 
  summarise(mean = mean(Socialization, na.rm = T))
# Mean socialization skills in ASD children was 77.36 and 101.70 in TD children

data %>% filter(VISIT == 1) %>% 
  count(Ethnicity)
# 2 African American, 1 Asian, 1 Bangladeshi, 1 Latino, 1 Lebanese, 57 White, 1 White/Asian, 2 White/Latino
```

### Chaging variable types

```{r Changing variable types}
df$Diagnosis <- as.factor(df$Diagnosis)
df$SUBJ <- as.factor(df$SUBJ)
df$Gender <- as.factor(df$Gender)
df$Ethnicity <- as.factor(df$Ethnicity)
df$VISIT <- as.numeric(df$VISIT)
```

### Visualising descriptive stats

```{r}
ggplot(df, aes(Gender, fill = Ethnicity)) + 
        geom_bar() +
        facet_grid(cols = vars(Diagnosis), margins = T) 
```

## Describe linguistic development (in terms of MLU over time) in TD and ASD children (as a function of group).


```{r}
ggplot(df, aes(x=VISIT, y = CHI_MLU, color = Diagnosis)) +
  geom_point() +
  geom_smooth(method=lm , color="black", se=TRUE) + 
  facet_grid(cols = vars(Diagnosis))
```


### Visualising the mothers' the childrens MLU over the 6 visits

```{r Animation of MLU per visit}
animation <- ggplot(df, aes(VISIT, CHI_MLU, color = Diagnosis)) +
  geom_point() +
  scale_x_log10() +
  theme_bw() +
  # gganimate specific bits:
  labs(title = 'VISIT: {frame_time}', x = 'VISIT', y = 'Child MLU') +
  transition_time(VISIT) +
  ease_aes('linear')

animate(animation, duration = 10, fps = 20, width = 200, height = 200, renderer = gifski_renderer())
anim_save("childmlu.gif")
```

```{r Visualising MLU the different visits, include = FALSE}
ggline(df,
       x = "VISIT",
       y = "CHI_MLU",
       col = "Diagnosis",
       add = c("mean_se", "dodge"),
       palette = "jco")
```

### Creating the first model for testing our hypothesis

```{r Creating model of MLU}
M1 <- lmer(CHI_MLU ~ Diagnosis * VISIT + (1 + VISIT | SUBJ), df, REML = F)
summary(M1)
performance::r2(M1)
```

## Describe how parental use of language (in terms of MLU) changes over time. What do you think is going on?

### Visualising the mothers' MLU as a function of visits

```{r}
ggline(df,
       x = "VISIT",
       y = "MOT_MLU",
       add = c("mean_se", "dodge"),
       palette = "jco")
```

### Making a model to describe the mothers' MLU over the different visits

```{r}
Model_MOT <- lmer(MOT_MLU ~ Diagnosis*VISIT + (1+VISIT|SUBJ), data = df, REML = FALSE)
summary(Model_MOT)
performance::r2(Model_MOT)
```

## Include individual differences in your model of language development (in children). Identify the best model.

### Making the final model and evaluating it

```{r Making final model, include = FALSE}
growth_m1 <- lmer(CHI_MLU ~  MullenRaw + (1 + VISIT | SUBJ), data = df, REML = FALSE)
growth_m1 <- lmer(CHI_MLU ~ Diagnosis * VISIT^2 + ExpressiveLangRaw + tokens_CHI + types_CHI + MOT_MLU + (1 + VISIT | SUBJ), data = df, REML = FALSE)
r2(growth_m1)
summary(growth_m1)
```

### Comparing the final model with less complex models

```{r Comparing final model to less complex ones}
M1 <- lmer(CHI_MLU ~ Diagnosis * VISIT^2 + (1 + VISIT | SUBJ), data = df, REML = FALSE)
M2 <- lmer(CHI_MLU ~ Diagnosis * VISIT^2 + (1 + VISIT | SUBJ), data = df, REML = FALSE)
M3 <- lmer(CHI_MLU ~ Diagnosis * VISIT^2 + tokens_CHI + (1 + VISIT | SUBJ), data = df, REML = FALSE)
M4 <- lmer(CHI_MLU ~ Diagnosis * VISIT^2 + tokens_CHI + types_CHI + (1 + VISIT | SUBJ), data = df, REML = FALSE)
M5 <- lmer(CHI_MLU ~ Diagnosis * VISIT^2 + tokens_CHI + types_CHI + MOT_MLU + (1 + VISIT | SUBJ), data = df, REML = FALSE)

anova(M1, M2, M3, M4, M5)
```

